---
title: "MSA 8265 Module 8 Assignment--DataToAction"
author: "Jennie Franco"
date: "5/14/2023"
output: pdf_document
---

# 1 Load libraries
```{r}
library(pacman)

# Work horse packages
library(tidyverse)
library(lubridate)

# theme_tq()
library(tidyquant)
library(ggplot2)

# Excel and CSV Files
library(readxl)
library(writexl)

p_load(janitor)
```

# 2 Importing Files
```{r}
auto <- read_csv("C:/Users/jenni/Documents/VillaNova University/05_Spring 2023/Data Management/Module 8/Metrics_Auto_Data_Year.csv")
```

# 3 Wrangling Data
```{r}
# Calculate Percent Change--Ford Motors Company
ford <- auto %>% 
    subset(companyName %in% c("Ford Motor Company")) %>% 
    group_by(dataItemName) %>% 
    mutate(pctChange = (totalValue/lag(totalValue) - 1) *100) %>% 
    mutate_at(7, round,2)
ford 

# Calculate Percent Change--General Motors Company
gm <- auto %>% 
    subset(companyName %in% c("General Motors Company")) %>% 
    group_by(dataItemName) %>% 
    mutate(pctChange = (totalValue/lag(totalValue) - 1) *100) %>% 
    mutate_at(7, round,2)
gm

# Bind the two dataframes
auto_wrangled <- rbind(ford,gm)
auto_wrangled
```


# 4 Visualizations
```{r}

# Step 1 - Manipulate

Revenue_by_year <-  auto_wrangled %>% 
    
    # Selecting columns to focus on
    
    select(companyName, dataItemName, pctChange, fiscalYear, totalValue) %>% 

    #Groupby and Summarize on year and performance metric
    group_by(fiscalYear, dataItemName) %>% 
    ungroup()

Revenue_by_year  
```



```{r fig.align="center", echo = FALSE}

# Step 2 - Visualize
Revenue_by_year %>% 
    
    # Setup canvas with year (x-axis) and pct-change (y-axis)
    ggplot(aes(x = fiscalYear, y = totalValue, fill=dataItemName)) + 
    
    # Geometries
    geom_col() +
    geom_smooth(method = "lm", se = FALSE) +
    geom_text(aes(label = pctChange), size = 3.1, vjust = -0.1, nudge_y = 4)+
    #geom_label(aes(label=pctChange), size = 3, nudge_x = -0.5, nudge_y = 5)+
    
    # Facet 
    facet_grid(companyName ~ dataItemName, scales = "free_y") +
    
    # Formatting
    theme_tq() +
    scale_fill_tq() +
    scale_x_continuous(breaks = seq(ceiling(min(Revenue_by_year$fiscalYear)),  floor(max(Revenue_by_year$fiscalYear)), by = 1)) +
    scale_y_continuous(labels = scales::comma) +
    labs(
        title = "Historical Performance of Ford & General Motors",
        subtitle = "Percent Change by Year",
        x = "",
        y = "Value",
        fill = "Performance Metric"
    )

#ggsave("HisotricalPerformance.png", width = 9, height = 6.5, path = "C:/Users/jenni/Documents/VillaNova University/05_Spring 2023/Data Management/Module 8/imgs" )
```

# 5 Correlations
```{r}
# Read in new data from Databricks code
ford_corr_data <- read_csv("C:/Users/jenni/Documents/VillaNova University/05_Spring 2023/Data Management/Module 8/Correlation Data/Ford_Data_Corr.csv")

gm_corr_data <- read_csv("C:/Users/jenni/Documents/VillaNova University/05_Spring 2023/Data Management/Module 8/Correlation Data/GM_Data_Corr.csv")

```

```{r}
fordResult = cor(ford_corr_data$tradeVolume, ford_corr_data$stockPrice, method = "pearson")
gmResult = cor(gm_corr_data$tradeVolume, gm_corr_data$stockPrice, method = "pearson")

#print the result
cat("Pearson correlation coefficient of Trade Volume & Stock Price for Ford is:", fordResult)

cat("Pearson correlation coefficient of Trade Volume & Stock Price for GM is:", gmResult)
```

Results:

* With a pearson correlation value of 0.45, there is a positive and moderate relationship between trade volume and stock price for Ford Motor Company.

* With a pearson correlation value of 0.19, there is a positive and weak relationship between trade volume and stock price for General Motors Company.


# 6 Additional Analyses
```{r}
# Read in new data from Databricks code
inv_data <- read_csv("C:/Users/jenni/Documents/VillaNova University/05_Spring 2023/Data Management/Module 8/Inventory_Auto_Data.csv")
```



```{r}
# Convert excel output from long to wide data format for analysis
library(tidyr)
inv_data_wide <- spread(inv_data, dataItemName, totalValue)

inv_data_wide <- inv_data_wide %>% clean_names()
```

```{r}
#Calculate Days Inventory Outstanding

inv_data_wide <- inv_data_wide %>% 
    mutate(days_inv_outstanding = ((inventory / cost_of_goods_sold_total)*365))

inv_data_wide
```

